---
title: "gamRR"
output: html_document
---

```{r setup, echo = FALSE, message=FALSE, warning=FALSE}
# load in useful packages
library(tidyverse)
library(forcats)
library(lubridate)
library(stringr)
library(ggplot2)
library(dplyr)
library(lmerTest)
library(lme4)
library(mgcv)
library(gridExtra)
library(patchwork)

# set default knitr chunks
knitr::opts_chunk$set(
  echo = FALSE,  # don't print the code chunk
  warning = FALSE,  # don't print warnings
  message = FALSE,  # don't print messages
  fig.width = 6,  # set default width of figures
  fig.height = 4,  # set default height of figures
  fig.align = "center",  # always align figure in center
  fig.pos = "H",  # always plot figure at the exact location of the code chunk
  cache = FALSE)  # don't cache results

reset_index <- function(df) {
  df <- cbind(index = rownames(df), df)
  rownames(df) <- 1:nrow(df)
  return(df)
}

listfunc.get <- function(ilst, str) {
  arg <- paste0("ilst$", str)
  return(eval(parse(text=arg)))
}
```

```{r load-data, echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE}
# note that the cache = TRUE chunk argument means that the results of this 
# chunk are saved so it doesn't need to be re-run every time you compile the pdf
# load the dates data
data_root <- "C:/wilbrechtlab/U19/data/RR/caching/gam/"
plot_path <- "C:/wilbrechtlab/U19/data/RR/plots/gam/"
source_file <- "all_data_gam_mat.csv"
psep = ".Platform$file.sep"

data <- read.table(paste0(data_root, source_file), sep=",", header=TRUE)
data <- mutate(data, animal=as.factor(animal),
               offer_prob=as.factor(offer_prob))
head(data)
```

```{r clean_data, cache=FALSE}
event_taus = list('tone_onset'=c(0, 1),
                  'acceptTJ'=c(-1, 1.5),
                  'rejectTJ'=c(-1, 1.5),
                  'quit'=c(-1.5, 1),
                  'rewardT'=c(0, 1.5),
                  'unrewardT'=c(0, 1.5),
                  'collection'=c(0, 1))

str.contains <- function(s, sub) {
  return(grepl(sub, s, fixed=TRUE))
}
strip.event_prefix <- function(s) {
  if (str.contains(s, 'next_') | str.contains(s, 'prev_')) {
    return(substr(s, 6, nchar(s)))
  } else {
    return(s)
  }
}

get_indic_vars <- function(df) {
  evcs <- colnames(df)[12:ncol(df)]
  ind_lst <- list()
  for (c in evcs) {
    ev_taus <- event_taus[[strip.event_prefix(c)]]
    vs <- df[,c]
    ind_lst[[paste0('I__', c)]] <- as.double((!is.na(vs)) & (vs>= ev_taus[1]) & (vs <= ev_taus[2]))
  }
  return(as.data.frame(do.call(cbind, ind_lst)))
}

source.unreg <- "all_data_gam_mat_unreg.csv"
data.raw <- read.table(paste0(data_root, source.unreg), sep=",", header=TRUE)
data <- cbind(data, get_indic_vars(data.raw))
data$tentry <- if_else(data$ACC==1, 
                       data$acceptTJ, data$rejectTJ)
```



```{r normalvis, cache=FALSE}
# d1_left <- data %>% filter((cell_type=='D1') & (hemi=='left'))
# ggplot(d1_left, aes(x=ZdF)) + geom_histogram()
# qqnorm(d1_left$ZdF, pch=1, frame=FALSE)
# qqline(d1_left$ZdF, col="steelblue", lwd=2)
```


## Problem Outline:

We have aligned neural and behavioral data from the RR task combining all animals, session, and `hemi`. We want to identify in `D1, A2A`, `left, right` hemisphere respectively, how are neural data represented in neural traces `ZdF`. We divide our goals into several steps

1.  First we focus on single session analysis
-   Each analysis we follow the following steps
  * fit GAM, check residual, record GCV score
  * plot partial dependence
  * `vis.gam` for tensor obj if needed
-   Assumption: $tau_0 = 0$ for each event. However GAM models may not be able to capture this due to identification problem. *A1*
-   *m.naive1* Naive GAM, with tone_onset:acc + acc/reject movement, and offer tone information (A1)
    -   compare without A1, using indicator variable instead. *m.naive0*
-   GAM with previous trial event: `prev_reject, prev_reward` etc. *m.prev*
- GAM with next trial event *m.next*
- GAM with all events *m.full*
2. Full session 
3. GAMM

```{r gam_examples, cache=FALSE}
d1_left <- data %>% filter((cell_type=='D1') & (hemi=='left')) %>%
        mutate(offer_prob=as.factor(offer_prob))
d1left.s <- d1_left %>% filter(animal=='RRM030' & session=='Day149')
ps_gam <- function (x) {
  return(s(x, bs='ps', k=30))
  
}
kspl = 50
m.naive1 <- gam(
  ZdF ~ s(tone_onset, bs='ps', k=kspl) +
    s(tone_onset, bs='ps', k=kspl, by=offer_prob) +
    s(tone_onset, bs='ps', k=kspl, by=ACC) +
    s(acceptTJ, bs='ps', k=kspl) +
    s(rejectTJ, bs='ps', k=kspl),
  data=d1left.s
)
# plot(m.naive1)
# summary(m.naive1)
# hist(d1left.s$ZdF)
```


```{r naive0, fig.height=6, fig.width=8, cache=FALSE}
d1left.s0 <- d1left.s
d1left.s0$tentry <- if_else(d1left.s0$ACC==1, 
                            d1left.s0$acceptTJ, d1left.s0$rejectTJ)
m.naive0 <- gam(
  ZdF ~ s(tone_onset, bs='ps', k=kspl) +
    s(tone_onset, bs='ps', k=kspl, by=offer_prob) +
    s(tone_onset, bs='ps', k=kspl, by=ACC) +
    s(tentry, bs='ps', k=kspl) +
    s(tentry, bs='ps', k=kspl, by=ACC),
  data=d1left.s0
)
pd<-plot(m.naive0, select=0)
dlst <- gam.pdgather(pd)
gam.ggplot(dlst)
```




```{r gamfuncs, fig.width=8, fig.height=6}

assign.ylab <- function(ylab) {
  colon <- ":"
  if (grepl(colon, ylab, fixed=TRUE)) {
    label <- strsplit(ylab, split=colon)[[1]][2]
  } else {
    label <- "base"
  }
  return(label)
}

gam.pdgather <- function(pd) {
  # This function unpacks GAM plotting data and gathers them into 
  # list of dataframes good for ggplot plotting
  dlst <- list()
  colon = ':'
  for (i in 1:length(pd)) {
    d <- pd[[i]]
    group <-d$xlab
    label <- assign.ylab(d$ylab)
    if (group %in% names(dlst)) {
      dlst[[group]][[label]] <- d$fit
    } else {
      xv <- d$x
      dlst[[group]] <- list('x'=xv)
      dlst[[group]][[label]] <- d$fit
    }
  }
  return(dlst)
}

gam.ggplot <- function(pd.list) {
  # takes in gathered gam plotting data and generates ggplots
  plist <- list()
  for(g in names(pd.list)) {
    df <- as.data.frame(do.call(cbind, pd.list[[g]]))
    colnames(df) <- names(pd.list[[g]])
    cns <- colnames(df)
    mut_cols <- cns[!(cns %in% c('x', 'base'))]
    if ('base' %in% cns) {
      df[,mut_cols] <- df[,mut_cols] + df[, 'base']
    }
    p <- gather(df, condition, ZdF, cns[cns != 'x']) %>%
      ggplot(aes(x=x, y=ZdF, group=condition, color=condition)) +
      geom_line() +
      labs(title=g)+
      theme(aspect.ratio = 0.8)
    plist[[length(plist) + 1]] <- p
  }
  ncol <- ceiling(sqrt(length(plist)))
  # grid.arrange(grobs=plist, ncol=ncol)
  # return(wrap_plots(plist) +
  #          theme(plot.margin = margin(20, 20, 20, 20),
  #          # Adjust overall plot size
  #          plot.background = element_rect(fill = "white", size = 1))
  #        )
  return(wrap_plots(plist, ncol=ncol))
}

run.gam <- function(data, formula, alias, ret_model=FALSE, plotpath=NULL) {
  # gam function that takes in a formula and a dataset, and obtain gcv, rsq, plots
  # alias organized as [roi]_[datascope]_[model]
  m <- gam(eval(parse(text=formula)),
    data=data)
  if (ret_model) {
    return(m)
  } else {
    dlst <- gam.pdgather(plot(m, select=0))
    ps <- gam.ggplot(dlst)
    fname <- paste0(alias, '.png')
    if (!is.null(plotpath)){
      len <- length(ps)
      cs <- ceiling(sqrt(len))
      rs <- ceiling(len/cs)
      ggsave(paste0(plot_path, fname), ps, height=4*rs, width=4*cs)
    }
    s <- summary(m)
    return(list('model'=alias, 'GCV'=s$sp.criterion[[1]], 'R2'=s$r.sq))
  }
}

format.m.res <- function(res.lst){
  return(paste0(res.lst[['alias']], 'GCV: ', round(res.lst[['GCV']], 4), 
                ' R2: ', round(res.lst[['R2']])))
}

pack.list.df <- function(named_lst) {
  result_df <- as.data.frame(named_lst)
  cvals <- list()
  cnames <- colnames(result_df)
  for (c in cnames){
    cvals[[length(cvals)+1]] <-do.call(rbind, result_df[,c])
  }
  result_df <- as.data.frame(do.call(cbind, cvals))
  colnames(result_df) <- cnames
  return(result_df)
}  

```

```{r prep_gam_data} 
rois <- list(
  'd1L'= filter(data, (cell_type=='D1') & (hemi=='left')),
  'd1R'=filter(data, (cell_type=='D1') & (hemi=='right')),
  'd2L'=filter(data, (cell_type=='A2A') & (hemi=='left')),
  'd2R'=filter(data, (cell_type=='A2A') & (hemi=='right'))
)

select.datascope <- function(d, scope, roi) {
  "Returns the data matching to scope and roi"
  if (grepl('d1', roi, fixed=TRUE)) {
    anim <- 'RRM030'
    sess <- 'Day159'
  } else {
    anim <- 'RRM032'
    sess <- 'Day135'
  }
  if (scope=='anim') {
    return(filter(d, animal==anim))
  } else if (scope=='sess') {
    return(filter(d, (animal==anim) & (session==sess)))
  } else { # full
    return(d)
  }
}
```

```{r naive_models}
m.formulas <- list()
kspl = 50 

formula = "ZdF ~ s(tone_onset, bs='ps', k=kspl) +
    s(tone_onset, bs='ps', k=kspl, by=offer_prob) +
    s(tone_onset, bs='ps', k=kspl, by=ACC) +
    I__acceptTJ + I__rejectTJ"

m.formulas[['naive_I']] = formula

formula = "ZdF ~ s(tone_onset, bs='ps', k=kspl) +
    s(tone_onset, bs='ps', k=kspl, by=offer_prob) +
    s(tone_onset, bs='ps', k=kspl, by=ACC) +
    s(neur_time, bs='ps', k=kspl, by=I__acceptTJ) + 
    s(neur_time, bs='ps', k=kspl, by=I__rejectTJ)"
m.formulas[['naive_It']] = formula

formula = "ZdF ~ s(tone_onset, bs='ps', k=kspl) +
    s(tone_onset, bs='ps', k=kspl, by=offer_prob) +
    s(tone_onset, bs='ps', k=kspl, by=ACC) +
    s(acceptTJ, bs='ps', k=kspl) +
    s(rejectTJ, bs='ps', k=kspl)"
m.formulas[['naive_g']] = formula

formula = "ZdF ~ s(tone_onset, bs='ps', k=kspl) +
    s(tone_onset, bs='ps', k=kspl, by=offer_prob) +
    s(tone_onset, bs='ps', k=kspl, by=ACC) +
    s(acceptTJ, bs='ps', k=kspl, by=I__acceptTJ) +
    s(rejectTJ, bs='ps', k=kspl, by=I__rejectTJ)"
m.formulas[['naive_Ig']] = formula

formula <- "ZdF ~ s(tone_onset, bs='ps', k=kspl) +
    s(tone_onset, bs='ps', k=kspl, by=offer_prob) +
    s(tone_onset, bs='ps', k=kspl, by=ACC) +
    s(tentry, bs='ps', k=kspl) +
    s(tentry, bs='ps', k=kspl, by=ACC)"
m.formulas[['naive_IgM']] = formula

rois <- list(
  'd1L'= filter(data, (cell_type=='D1') & (hemi=='left')),
  'd1R'=filter(data, (cell_type=='D1') & (hemi=='right')),
  'd2L'=filter(data, (cell_type=='A2A') & (hemi=='left')),
  'd2R'=filter(data, (cell_type=='A2A') & (hemi=='right'))
)

select.datascope <- function(d, scope, roi) {
  "Returns the data matching to scope and roi"
  if (grepl('d1', roi, fixed=TRUE)) {
    anim <- 'RRM030'
    sess <- 'Day159'
  } else {
    anim <- 'RRM032'
    sess <- 'Day135'
  }
  if (scope=='anim') {
    return(filter(d, animal==anim))
  } else if (scope=='sess') {
    return(filter(d, (animal==anim) & (session==sess)))
  } else { # full
    return(d)
  }
}

gam.results <- list()

for (roi in names(rois)) {
  for(scope in c('sess', 'anim', 'full')) {
    d <- select.datascope(rois[[roi]], scope, roi)
    for (mname in names(m.formulas)) {
      alias <- paste0(roi, '_', scope, '_', mname)
      res<-run.gam(d, m.formulas[[mname]], alias, ret_model = FALSE, plotpath=plot_path)
      gam.results[[length(gam.results)+1]] = res
    }
  }
}
gam.results <- do.call(rbind, gam.results)
result_df <- pack.list.df(gam.results) %>%
  separate(model, c('roi', 'datascope', 'model_class', 'model'), sep='_') %>%
  mutate(model=paste0(model_class, '_', model)) %>%
  arrange(roi, datascope, GCV)
result_df
```

```{r rungam}
m.formulas <- list()
kspl = 50 

# formula = "ZdF ~ s(tone_onset, bs='ps', k=kspl) +
#     s(tone_onset, bs='ps', k=kspl, by=offer_prob) +
#     s(tone_onset, bs='ps', k=kspl, by=ACC) +
#     I__acceptTJ + I__rejectTJ"
# 
# m.formulas[['naive_I']] = formula
# 
# formula = "ZdF ~ s(tone_onset, bs='ps', k=kspl) +
#     s(tone_onset, bs='ps', k=kspl, by=offer_prob) +
#     s(tone_onset, bs='ps', k=kspl, by=ACC) +
#     s(neur_time, bs='ps', k=kspl, by=I__acceptTJ) + 
#     s(neur_time, bs='ps', k=kspl, by=I__rejectTJ)"
# m.formulas[['naive_It']] = formula
# 
# formula = "ZdF ~ s(tone_onset, bs='ps', k=kspl) +
#     s(tone_onset, bs='ps', k=kspl, by=offer_prob) +
#     s(tone_onset, bs='ps', k=kspl, by=ACC) +
#     s(acceptTJ, bs='ps', k=kspl) +
#     s(rejectTJ, bs='ps', k=kspl)"
# m.formulas[['naive_g']] = formula
# 
# formula = "ZdF ~ s(tone_onset, bs='ps', k=kspl) +
#     s(tone_onset, bs='ps', k=kspl, by=offer_prob) +
#     s(tone_onset, bs='ps', k=kspl, by=ACC) +
#     s(acceptTJ, bs='ps', k=kspl, by=I__acceptTJ) +
#     s(rejectTJ, bs='ps', k=kspl, by=I__rejectTJ)"
# m.formulas[['naive_Ig']] = formula
# 
# formula <- "ZdF ~ s(tone_onset, bs='ps', k=kspl) +
#     s(tone_onset, bs='ps', k=kspl, by=offer_prob) +
#     s(tone_onset, bs='ps', k=kspl, by=ACC) +
#     s(tentry, bs='ps', k=kspl) +
#     s(tentry, bs='ps', k=kspl, by=ACC)"
# m.formulas[['naive_IgM']] = formula

# prev_tone_onset, prev_acceptTJ, prev_rejectTJ, prev_quit, prev_rewardT, prev_unrewardT
# Ig, g, It

formula <- "ZdF ~ s(tone_onset, bs='ps', k=kspl) +
    s(tone_onset, bs='ps', k=kspl, by=offer_prob) +
    s(tone_onset, bs='ps', k=kspl, by=ACC) +
    s(acceptTJ, bs='ps', k=kspl, by=I__acceptTJ) +
    s(rejectTJ, bs='ps', k=kspl, by=I__rejectTJ) +
    s(neur_time, bs='ps', k=kspl, by=I__prev_tone_onset) +
    s(neur_time, bs='ps', k=kspl, by=I__prev_acceptTJ) +
    s(neur_time, bs='ps', k=kspl, by=I__prev_rejectTJ) +
    s(neur_time, bs='ps', k=kspl, by=I__prev_quit) + 
    s(neur_time, bs='ps', k=kspl, by=I__prev_rewardT) + 
    s(neur_time, bs='ps', k=kspl, by=I__prev_unrewardT)
"
m.formulas[['prev_It']] = formula

formula <- "ZdF ~ s(tone_onset, bs='ps', k=kspl) +
    s(tone_onset, bs='ps', k=kspl, by=offer_prob) +
    s(tone_onset, bs='ps', k=kspl, by=ACC) +
    s(acceptTJ, bs='ps', k=kspl, by=I__acceptTJ) +
    s(rejectTJ, bs='ps', k=kspl, by=I__rejectTJ) +
    s(prev_tone_onset, bs='ps', k=kspl) +
    s(prev_acceptTJ, bs='ps', k=kspl) +
    s(prev_rejectTJ, bs='ps', k=kspl) +
    s(prev_quit, bs='ps', k=kspl) + 
    s(prev_rewardT, bs='ps', k=kspl) + 
    s(prev_unrewardT, bs='ps', k=kspl)"

m.formulas[['prev_g']] = formula

formula <- "ZdF ~ s(tone_onset, bs='ps', k=kspl) +
    s(tone_onset, bs='ps', k=kspl, by=offer_prob) +
    s(tone_onset, bs='ps', k=kspl, by=ACC) +
    s(acceptTJ, bs='ps', k=kspl, by=I__acceptTJ) +
    s(rejectTJ, bs='ps', k=kspl, by=I__rejectTJ) +
    s(prev_tone_onset, bs='ps', k=kspl, by=I__prev_tone_onset) +
    s(prev_acceptTJ, bs='ps', k=kspl, by=I__prev_acceptTJ) +
    s(prev_rejectTJ, bs='ps', k=kspl, by=I__prev_rejectTJ) +
    s(prev_quit, bs='ps', k=kspl, by=I__prev_quit) + 
    s(prev_rewardT, bs='ps', k=kspl, by=I__prev_rewardT) + 
    s(prev_unrewardT, bs='ps', k=kspl, by=I__prev_unrewardT)"
m.formulas[['prev_Ig']] = formula

# next_tone_onset, next_acceptTJ, next_rejectTJ, next_quit

formula <- "ZdF ~ s(tone_onset, bs='ps', k=kspl) +
    s(tone_onset, bs='ps', k=kspl, by=offer_prob) +
    s(tone_onset, bs='ps', k=kspl, by=ACC) +
    s(acceptTJ, bs='ps', k=kspl, by=I__acceptTJ) +
    s(rejectTJ, bs='ps', k=kspl, by=I__rejectTJ) +
    s(neur_time, bs='ps', k=kspl, by=I__next_tone_onset) +
    s(neur_time, bs='ps', k=kspl, by=I__next_acceptTJ) +
    s(neur_time, bs='ps', k=kspl, by=I__next_rejectTJ) +
    s(neur_time, bs='ps', k=kspl, by=I__next_quit)"
m.formulas[['next_It']] = formula

formula <- "ZdF ~ s(tone_onset, bs='ps', k=kspl) +
    s(tone_onset, bs='ps', k=kspl, by=offer_prob) +
    s(tone_onset, bs='ps', k=kspl, by=ACC) +
    s(acceptTJ, bs='ps', k=kspl, by=I__acceptTJ) +
    s(rejectTJ, bs='ps', k=kspl, by=I__rejectTJ) +
    s(next_tone_onset, bs='ps', k=kspl) +
    s(next_acceptTJ, bs='ps', k=kspl) +
    s(next_rejectTJ, bs='ps', k=kspl) +
    s(next_quit, bs='ps', k=kspl)"

m.formulas[['next_g']] = formula

formula <- "ZdF ~ s(tone_onset, bs='ps', k=kspl) +
    s(tone_onset, bs='ps', k=kspl, by=offer_prob) +
    s(tone_onset, bs='ps', k=kspl, by=ACC) +
    s(acceptTJ, bs='ps', k=kspl, by=I__acceptTJ) +
    s(rejectTJ, bs='ps', k=kspl, by=I__rejectTJ) +
    s(next_tone_onset, bs='ps', k=kspl, by=I__next_tone_onset) +
    s(next_acceptTJ, bs='ps', k=kspl, by=I__next_acceptTJ) +
    s(next_rejectTJ, bs='ps', k=kspl, by=I__next_rejectTJ) +
    s(next_quit, bs='ps', k=kspl, by=I__next_quit)"
m.formulas[['next_Ig']] = formula

gam.results <- list()

for (roi in names(rois)) {
  for(scope in c('sess', 'anim', 'full')) {
    d <- select.datascope(rois[[roi]], scope, roi)
    for (mname in names(m.formulas)) {
      alias <- paste0(roi, '_', scope, '_', mname)
      res<-run.gam(d, m.formulas[[mname]], alias, ret_model = FALSE, plotpath=plot_path)
      gam.results[[length(gam.results)+1]] = res
    }
  }
}
gam.results <- do.call(rbind, gam.results)
result_df <- pack.list.df(gam.results) %>%
  separate(model, c('roi', 'datascope', 'model_class', 'model'), sep='_') %>%
  mutate(model=paste0(model_class, '_', model)) %>%
  arrange(roi, datascope, GCV)
result_df
```

```{r more_models}
# write.csv(result_df, paste0(data_root, 'gam_result_prevnext.csv'))

m.formulas <- list()
kspl = 50 

formula <- "ZdF ~ s(tone_onset, bs='ps', k=kspl) +
    s(tone_onset, bs='ps', k=kspl, by=offer_prob) +
    s(tone_onset, bs='ps', k=kspl, by=ACC) +
    s(acceptTJ, bs='ps', k=kspl, by=I__acceptTJ) +
    s(rejectTJ, bs='ps', k=kspl, by=I__rejectTJ) +
    s(neur_time, bs='ps', k=kspl, by=I__prev_tone_onset) +
    s(neur_time, bs='ps', k=kspl, by=I__prev_acceptTJ) +
    s(neur_time, bs='ps', k=kspl, by=I__prev_rejectTJ) +
    s(neur_time, bs='ps', k=kspl, by=I__prev_quit) + 
    s(neur_time, bs='ps', k=kspl, by=I__prev_rewardT) + 
    s(neur_time, bs='ps', k=kspl, by=I__prev_unrewardT) + 
    s(neur_time, bs='ps', k=kspl, by=I__next_tone_onset) +
    s(neur_time, bs='ps', k=kspl, by=I__next_acceptTJ) +
    s(neur_time, bs='ps', k=kspl, by=I__next_rejectTJ) +
    s(neur_time, bs='ps', k=kspl, by=I__next_quit)"
m.formulas[['combined_It']] = formula

formula <- "ZdF ~ s(tone_onset, bs='ps', k=kspl) +
    s(tone_onset, bs='ps', k=kspl, by=offer_prob) +
    s(tone_onset, bs='ps', k=kspl, by=ACC) +
    s(acceptTJ, bs='ps', k=kspl, by=I__acceptTJ) +
    s(rejectTJ, bs='ps', k=kspl, by=I__rejectTJ) +
    s(prev_tone_onset, bs='ps', k=kspl) +
    s(prev_acceptTJ, bs='ps', k=kspl) +
    s(prev_rejectTJ, bs='ps', k=kspl) +
    s(prev_quit, bs='ps', k=kspl) + 
    s(prev_rewardT, bs='ps', k=kspl) + 
    s(prev_unrewardT, bs='ps', k=kspl)+
    s(next_tone_onset, bs='ps', k=kspl) +
    s(next_acceptTJ, bs='ps', k=kspl) +
    s(next_rejectTJ, bs='ps', k=kspl) +
    s(next_quit, bs='ps', k=kspl)
"
m.formulas[['combined_g']] = formula

formula <- "ZdF ~ s(tone_onset, bs='ps', k=kspl) +
    s(tone_onset, bs='ps', k=kspl, by=offer_prob) +
    s(tone_onset, bs='ps', k=kspl, by=ACC) +
    s(acceptTJ, bs='ps', k=kspl, by=I__acceptTJ) +
    s(rejectTJ, bs='ps', k=kspl, by=I__rejectTJ) +
    s(prev_tone_onset, bs='ps', k=kspl, by=I__prev_tone_onset) +
    s(prev_acceptTJ, bs='ps', k=kspl, by=I__prev_acceptTJ) +
    s(prev_rejectTJ, bs='ps', k=kspl, by=I__prev_rejectTJ) +
    s(prev_quit, bs='ps', k=kspl, by=I__prev_quit) + 
    s(prev_rewardT, bs='ps', k=kspl, by=I__prev_rewardT) + 
    s(prev_unrewardT, bs='ps', k=kspl, by=I__prev_unrewardT) +
    s(next_tone_onset, bs='ps', k=kspl, by=I__next_tone_onset) +
    s(next_acceptTJ, bs='ps', k=kspl, by=I__next_acceptTJ) +
    s(next_rejectTJ, bs='ps', k=kspl, by=I__next_rejectTJ) +
    s(next_quit, bs='ps', k=kspl, by=I__next_quit)
"
m.formulas[['combined_Ig']] = formula


library(foreach)
library(doParallel)
cores= detectCores()
print(paste0(cores, ' cores'))
cl <- makeCluster(cores[1]-2)
registerDoParallel(cl)
d.scopes <- c('sess', 'anim', 'full')
comb.groups <-expand.grid(names(rois), d.scopes)

gam.results.complex <- foreach(i=1:length(comb.groups), 
                               .combine=rbind,
                               .packages = c('tidyverse', 'mgcv', 
                                             'patchwork'),
                               .export=c('kspl', 'comb.groups')) %dopar% {
  roi <- as.character(comb.groups[i, 1])
  scope <- as.character(comb.groups[i, 2])
  # rlist <- c(roi, scope)
  d <- select.datascope(rois[[roi]], scope, roi)
  gam.rs <- list()
  for (mname in names(m.formulas)) {
    alias <- paste0(roi, '_', scope, '_', mname)
    res<-run.gam(d, m.formulas[[mname]], alias, ret_model = FALSE, plotpath=plot_path)
    gam.rs[[length(gam.rs)+1]] = res
  }
  rlist <- do.call(rbind, gam.rs)
  return(rlist)
}

result_df.complex <- pack.list.df(gam.results.complex) %>%
  separate(model, c('roi', 'datascope', 'model_class', 'model'), sep='_') %>%
  mutate(model=paste0(model_class, '_', model)) %>%
  arrange(roi, datascope, GCV)
write.csv(result_df.complex, paste0(data_root, 'gam_result_combined.csv'))
result_df.complex
```

```{r example, fig.height=40, fig.width=40}
kspl <- 50
formula <- "ZdF ~ s(tone_onset, bs='ps', k=kspl) +
    s(tone_onset, bs='ps', k=kspl, by=offer_prob) +
    s(tone_onset, bs='ps', k=kspl, by=ACC) +
    s(acceptTJ, bs='ps', k=kspl, by=I__acceptTJ) +
    s(rejectTJ, bs='ps', k=kspl, by=I__rejectTJ) +
    s(prev_tone_onset, bs='ps', k=kspl) +
    s(prev_acceptTJ, bs='ps', k=kspl) +
    s(prev_rejectTJ, bs='ps', k=kspl) +
    s(prev_quit, bs='ps', k=kspl) + 
    s(prev_rewardT, bs='ps', k=kspl) + 
    s(prev_unrewardT, bs='ps', k=kspl)+
    s(next_tone_onset, bs='ps', k=kspl) +
    s(next_acceptTJ, bs='ps', k=kspl) +
    s(next_rejectTJ, bs='ps', k=kspl) +
    s(next_quit, bs='ps', k=kspl)
"
roi <- 'd1L'
scope <- 'sess'
d <- select.datascope(rois[[roi]], scope, roi)
mname <- 'combined_g'
alias <- paste0(roi, '_', scope, '_', mname)
m<-run.gam(d, formula, alias, ret_model = TRUE)
dlst <- gam.pdgather(plot(m, select=0))
ps <- gam.ggplot(dlst)
ps

```

```{r}
mgcv::gam.check(m)
```



```{r diagnose}
result.cls <- c('gam_result_combined.csv',
                'gam_result_naive.csv',
                'gam_result_prevnext.csv')
df.lsts <-list()
for (rcls in result.cls) {
  df.lsts[[length(df.lsts)+1]] <- read.csv(paste0(data_root, rcls))
}
result_df.meta <- do.call(rbind, df.lsts) %>% arrange(roi, datascope, GCV) %>% 
  filter(datascope=='sess')
result_df.meta             
```

```{r}
kspl <- 50
# formula <- "ZdF ~ s(tone_onset, bs='ps', k=kspl) +
#     s(tone_onset, bs='ps', k=kspl, by=offer_prob) +
#     s(tone_onset, bs='ps', k=kspl, by=ACC) +
#     s(acceptTJ, bs='ps', k=kspl)+
#     s(acceptTJ, bs='ps', k=kspl, by=as.factor(acceptTJ==0)) +
#     s(rejectTJ, bs='ps', k=kspl)
# "
formula <- "ZdF ~ s(tone_onset, bs='ps', k=kspl) +
    s(tone_onset, bs='ps', k=kspl, by=offer_prob) +
    s(tone_onset, bs='ps', k=kspl, by=ACC) +
    s(acceptTJ, bs='ps', k=kspl)+
    s(acceptTJ, bs='ps', k=kspl, by=Iaccept0) +
    s(rejectTJ, bs='ps', k=kspl)
"
roi <- 'd1L'
scope <- 'sess'
d <- select.datascope(rois[[roi]], scope, roi)
mname <- 'combined_g'
alias <- paste0(roi, '_', scope, '_', mname)
m<-run.gam(d%>%mutate(Iaccept0=as.double(acceptTJ==0)), formula, alias, ret_model = TRUE)
dlst <- gam.pdgather(plot(m, select=0))
ps <- gam.ggplot(dlst) # compare indicator versus 
ps
```

```{r practice}
library(mgcv)
# Function to choose lambda using GCV
choose_lambda <- function(x, y, h_length=20, k=10, 
                         lambda_seq=exp(seq(-10, 10, length=100))) {
  # Setup initial model
  tau <- seq(0, 1, length.out=h_length)
  bp <- setup_basis_penalty(tau, k)
  Z <- compute_Z(x, bp$X)
  w <- rep(1, length(y))
  b <- rep(0, ncol(Z))
  
  # Function to compute GCV for given lambda
  compute_gcv <- function(lambda) {
    fit <- pcls(y, Z, w, bp$S, lambda, b)
    yhat <- Z %*% fit
    n <- length(y)
    # Effective degrees of freedom
    A <- solve(crossprod(Z) + lambda * bp$S)
    df <- sum(diag(Z %*% A %*% t(Z)))
    # GCV score
    gcv <- mean((y - yhat)^2) / (1 - df/n)^2
    return(gcv)
  }
  
  # Find best lambda
  gcv_scores <- sapply(lambda_seq, compute_gcv)
  best_lambda <- lambda_seq[which.min(gcv_scores)]
  
  return(best_lambda)
}

# Function to create basis and penalty
setup_basis_penalty <- function(tau, k=10) {
  # Create smooth specification
  sm <- smoothCon(s(tau, k=k, bs="cr", m=2), 
                  data=data.frame(tau=tau))[[1]]
  return(sm)
}

# Function to compute convolution matrix Z
compute_Z <- function(x, basis_matrix) {
  n <- length(x)
  k <- ncol(basis_matrix)
  Z <- matrix(0, nrow=n, ncol=k)
  
  # For each basis function
  for(j in 1:k) {
    phi_j <- basis_matrix[,j]
    # Compute convolution with x
    for(t in 1:n) {
      max_lag <- min(t, length(phi_j))
      Z[t,j] <- sum(x[t - seq_len(max_lag) + 1] * phi_j[seq_len(max_lag)])
    }
  }
  return(Z)
}

# Function to fit model using pcls
fit_conv_pcls <- function(x, y, h_length=20, k=10, sp=1) {
  # Setup basis and penalty
  # print(sp)
  tau <- seq(0, 1, length.out=h_length)
  sm <- setup_basis_penalty(tau, k)
  
  # Compute Z matrix (our design matrix)
  Z <- compute_Z(x, sm$X)
  # Z <- scale(Z)
  
  # Create the pcls input list
  # M <- list(
  #   y = y,                    # Response
  #   w = rep(1, length(y)),   # Weights
  #   X = Z,                    # Design matrix
  #   C = matrix(0, 1, ncol(Z)),     # No equality constraints
  #   S = sm$S,                # List of penalty matrices
  #   off = 0,            # Offset for penalties
  #   sp = sp,                 # Smoothing parameter
  #   p = rnorm(ncol(Z), sd=0.01),     # Initial parameter values
  #   Ain = matrix(0, 1, ncol(Z)),   # No inequality constraints
  #   bin = rep(-10000, 1)         # No inequality constraints
  # )
  # Fit using pcls
  #fit <- pcls(M)
  fit<-mgcv::mgcv(y,Z,sp,sm$S,off=0)
  
  return(list(
    coef = fit$b,
    basis = sm$X,
    tau = tau,
    Z = Z
  ))
}

# Simulation example
set.seed(123)
n <- 100
h_length <- 20

# Generate data
x <- rnorm(n)
tau <- seq(0, 1, length.out=h_length)
h_true <- exp(-3*tau) * sin(2*pi*5*tau)

# Create convolution matrix for data generation
X <- matrix(0, nrow=n, ncol=h_length)
for(i in 1:n) {
  for(j in 1:min(i, h_length)) {
    X[i,j] <- x[i-j+1]
  }
}
y <- X %*% h_true + rnorm(n, sd=0.1)

# Choose lambda using GCV
# best_lambda <- choose_lambda(x, y, h_length)
# fit <- pcls(y, w, w, bp$S, lambda, b)
# print(paste("Best lambda:", best_lambda))

# Fit model
fit <- fit_conv_pcls(x, y, h_length, k=10)

# Reconstruct h(tau)
h_est <- fit$basis %*% fit$coef

# Plot results
par(mfrow=c(1,2))
#Plot true vs estimated impulse response
plot(tau, h_true, type='l', col='blue',
     main='True vs Estimated Impulse Response',
     xlab='tau', ylab='h(tau)')
lines(tau, h_est, col='red', lty=2)
legend('topright', legend=c('True', 'Estimated'),
       col=c('blue', 'red'), lty=1:2)

# Plot observed vs fitted
y_fitted <- fit$Z %*% fit$coef
plot(y, y_fitted,
     main='Observed vs Fitted Values',
     xlab='Observed', ylab='Fitted')
abline(0, 1, col='red', lty=2)
```
