---
title: "RR ArchT GLMER"
author: "Albert Qü"
date: "`r format(Sys.time(), '%B %d, %Y')`"
header-includes:
   - \usepackage{float}
output: 
  pdf_document:
    number_sections: true
---
check tutorials: 
https://ourcodingclub.github.io/tutorials/mixed-models/


```{r setup, echo = FALSE, message=FALSE, warning=FALSE}
# load in useful packages
library(tidyverse)
library(forcats)
library(lubridate)
library(stringr)
library(ggplot2)
library(dplyr)
# set default knitr chunks
knitr::opts_chunk$set(
  echo = FALSE,  # don't print the code chunk
  warning = FALSE,  # don't print warnings
  message = FALSE,  # don't print messages
  fig.width = 6,  # set default width of figures
  fig.height = 4,  # set default height of figures
  fig.align = "center",  # always align figure in center
  fig.pos = "H",  # always plot figure at the exact location of the code chunk
  cache = FALSE)  # don't cache results

reset_index <- function(df) {
  df <- cbind(index = rownames(df), df)
  rownames(df) <- 1:nrow(df)
  return(df)
}
```


```{r load-data, echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE}
# note that the cache = TRUE chunk argument means that the results of this 
# chunk are saved so it doesn't need to be re-run every time you compile the pdf
# load the dates data
data_root <- "D:/U19/data/RR/"
source_file <- 'rrArchT_glmer_R.csv'
psep = ".Platform$file.sep"

data <- read.table(paste0(data_root, source_file), sep=",", header=TRUE)
#data <- mutate(data, across(c("stimType", "restaurant"), as.factor)) 
data <- mutate(data, animal=as.factor(animal), accept=as.double(accept),
               stimType=relevel(as.factor(stimType), "nostim"),
               restaurant=relevel(as.factor(restaurant), "2")) 
head(data)
```
```{r}
data %>% 
  select(animal, restaurant, tone_prob, stimType, accept, plugged_in, quit_time) %>% 
  filter(stimType %in% c('stim', 'nostim'), animal=="RRM039", plugged_in=="L", !is.na(quit_time))
```


## Data consistency check with Python Psychometric Curve

```{r stim_only_44, echo = TRUE, message=FALSE, warning=FALSE}
data.44.L <- data %>% 
  select(animal, restaurant, tone_prob, stimType, accept, plugged_in, quit_time) %>% 
  filter(stimType %in% c('stim', 'nostim'), animal=="RRM044", plugged_in=="L", is.na(quit_time))
glm.44.L <- glm(accept ~ restaurant:stimType + tone_prob:stimType + stimType + restaurant + tone_prob, family=binomial(), data=data.44.L)
summary(glm.44.L)

data.44.R <- data %>% 
  select(animal, restaurant, tone_prob, stimType, accept, plugged_in, quit_time) %>% 
  filter(stimType %in% c('stim', 'nostim'), animal=="RRM044", plugged_in=="R", is.na(quit_time))
glm.44.R <- glm(accept ~ restaurant:stimType + tone_prob:stimType + stimType + restaurant + tone_prob, family=binomial(), data=data.44.R)
summary(glm.44.R)

```

```{r}
data.44.L %>% filter(stimType=='nostim') %>% 
  group_by(restaurant, tone_prob) %>%
  summarise(accept_rate=mean(accept))
```

## Formal p value summary multiple animal/hemisphere (accept decision)

```{r stim_only_summary, echo = TRUE, message=FALSE, warning=FALSE}
df_list = list()
i = 1
for (animal_num in c(39:40, 42:45)) {
  animal_arg <- sprintf("RRM0%d", animal_num)
  for (hemi_arg in c("L", "R")) {
    data.ah <- data %>% 
    select(animal, restaurant, tone_prob, stimType, accept, plugged_in, quit_time) %>% 
    filter(stimType %in% c('stim', 'nostim'), animal==animal_arg, plugged_in==hemi_arg, is.na(quit_time))
    glm.ah <- glm(accept ~ restaurant:stimType + tone_prob:stimType + stimType + restaurant + tone_prob, family=binomial(), data=data.ah)
    print(paste(animal_arg, hemi_arg))
    print(summary(glm.ah))
    
    df <- reset_index(as.data.frame(coef(summary(glm.ah))[, c(1, 2, 4)]))
    df$animal = animal_arg
    df$hemi = hemi_arg
    df_list[[i]] = df
    i = i + 1
  }
}

accept_p_df <- do.call(rbind, df_list)
names(accept_p_df)[3:4] = c("stdE", "p_val")
accept_p_df %>% filter(p_val <= 0.05, grepl("stim", index))
```

## Formal p value summary multiple animal/hemisphere (decision time)

no effect of stim on decision time

summary: robust tone_prob effect on movement time, suggesting decision time positively correlated with tone probability

```{r stim_only_dt_summary, echo = TRUE, message=FALSE, warning=FALSE}
df_list = list()
i = 1
for (animal_num in c(39:40, 42:45)) {
  animal_arg <- sprintf("RRM0%d", animal_num)
  for (hemi_arg in c("L", "R")) {
    data.ah <- data %>% 
    select(animal, restaurant, tone_prob, stimType, decision_time, plugged_in, quit_time) %>% 
    filter(stimType %in% c('stim', 'nostim'), animal==animal_arg, plugged_in==hemi_arg)
    glm.ah <- glm(decision_time ~ stimType + restaurant + tone_prob, family=Gamma, data=data.ah)
    print(paste(animal_arg, hemi_arg))
    print(summary(glm.ah))
    
    df <- reset_index(as.data.frame(coef(summary(glm.ah))[, c(1, 2, 4)]))
    df$animal = animal_arg
    df$hemi = hemi_arg
    df_list[[i]] = df
    i = i + 1
  }
}

dt_p_df <- do.call(rbind, df_list)
names(dt_p_df)[3:4] = c("stdE", "p_val")
dt_p_df %>% filter(p_val <= 0.05, grepl("tone", index))
```


```{r stim_only_interaction_dt, echo = TRUE, message=FALSE, warning=FALSE}
df_list = list()
i = 1
for (animal_num in c(39:40, 42:45)) {
  animal_arg <- sprintf("RRM0%d", animal_num)
  for (hemi_arg in c("L", "R")) {
    data.ah <- data %>% 
    select(animal, restaurant, tone_prob, stimType, decision_time, plugged_in, quit_time) %>% 
    filter(stimType %in% c('stim', 'nostim'), animal==animal_arg, plugged_in==hemi_arg)
    glm.ah <- glm(decision_time ~ restaurant:stimType + tone_prob:stimType + stimType + restaurant + tone_prob, family=Gamma, data=data.ah)
    print(paste(animal_arg, hemi_arg))
    print(summary(glm.ah))
    
    df <- reset_index(as.data.frame(coef(summary(glm.ah))[, c(1, 2, 4)]))
    df$animal = animal_arg
    df$hemi = hemi_arg
    df_list[[i]] = df
    i = i + 1
  }
}

dt_p_df <- do.call(rbind, df_list)
names(dt_p_df)[3:4] = c("stdE", "p_val")
dt_p_df %>% filter(p_val <= 0.05, grepl("tone", index))
```


```{r stim_only_gaussian_dt, echo = TRUE, message=FALSE, warning=FALSE}
df_list = list()
i = 1
for (animal_num in c(39:40, 42:45)) {
  animal_arg <- sprintf("RRM0%d", animal_num)
  for (hemi_arg in c("L", "R")) {
    data.ah <- data %>% 
    select(animal, restaurant, tone_prob, stimType, decision_time, plugged_in, quit_time) %>% 
    filter(stimType %in% c('stim', 'nostim'), animal==animal_arg, plugged_in==hemi_arg)
    glm.ah <- glm(decision_time ~ restaurant:stimType + tone_prob:stimType + stimType + restaurant + tone_prob, family=gaussian(), data=data.ah)
    print(paste(animal_arg, hemi_arg))
    print(summary(glm.ah))
    
    df <- reset_index(as.data.frame(coef(summary(glm.ah))[, c(1, 2, 4)]))
    df$animal = animal_arg
    df$hemi = hemi_arg
    df_list[[i]] = df
    i = i + 1
  }
}

dt_p_df <- do.call(rbind, df_list)
names(dt_p_df)[3:4] = c("stdE", "p_val")
dt_p_df %>% filter(p_val <= 0.05, grepl("tone", index))
```


```{r stim_only_gaussian_dt_nointer, echo = TRUE, message=FALSE, warning=FALSE}
df_list = list()
i = 1
for (animal_num in c(39:40, 42:45)) {
  animal_arg <- sprintf("RRM0%d", animal_num)
  for (hemi_arg in c("L", "R")) {
    data.ah <- data %>% 
    select(animal, restaurant, tone_prob, stimType, decision_time, plugged_in, quit_time) %>% 
    filter(stimType %in% c('stim', 'nostim'), animal==animal_arg, plugged_in==hemi_arg)
    glm.ah <- glm(decision_time ~ stimType + restaurant + tone_prob, family=gaussian(), data=data.ah)
    print(paste(animal_arg, hemi_arg))
    print(summary(glm.ah))
    
    df <- reset_index(as.data.frame(coef(summary(glm.ah))[, c(1, 2, 4)]))
    df$animal = animal_arg
    df$hemi = hemi_arg
    df_list[[i]] = df
    i = i + 1
  }
}

dt_p_df <- do.call(rbind, df_list)
names(dt_p_df)[3:4] = c("stdE", "p_val")
dt_p_df %>% filter(p_val <= 0.05, grepl("stim", index))
```



GLM gaussian and lm yield similar results

```{r, fig.width=10,fig.height=11}
data %>% 
  select(animal, restaurant, tone_prob, stimType, decision_time, plugged_in, quit_time) %>% 
  filter(stimType %in% c('stim', 'nostim')) %>% 
  ggplot(aes(x=tone_prob, y=decision_time, color=stimType)) +
  geom_smooth(method='lm') + facet_wrap(animal ~ plugged_in)


d39L <- data %>% 
  select(animal, restaurant, tone_prob, stimType, decision_time, plugged_in, quit_time) %>% 
  filter(stimType== 'nostim', animal=='RRM039', plugged_in=='L')  

glm(decision_time ~ restaurant + tone_prob, data=d39L, family=Gamma) %>% summary()

glm(decision_time ~ restaurant + tone_prob, data=d39L, family=gaussian()) %>% summary()
```



## EDA: stimulation on quit_time
```{r, fig.width=15,fig.height=11}
data %>% 
  select(animal, restaurant, tone_prob, stimType, decision_time, plugged_in, quit_time) %>% 
  mutate(tone_prob=as.factor(tone_prob)) %>%
  filter(stimType %in% c('stim', 'nostim'), !is.na(quit_time)) %>%
  ggplot(aes(x=tone_prob, y=quit_time, color=stimType)) +
  geom_boxplot() + facet_wrap(animal ~ plugged_in)


df_list = list()
i = 1
for (animal_num in c(39:40, 42:45)) {
  animal_arg <- sprintf("RRM0%d", animal_num)
  for (hemi_arg in c("L", "R")) {
    data.ah <- data %>% 
    select(animal, restaurant, tone_prob, stimType, decision_time, plugged_in, quit_time) %>% 
    filter(stimType %in% c('stim', 'nostim'), !is.na(quit_time), animal==animal_arg, plugged_in==hemi_arg)
    glm.ah <- glm(quit_time ~ stimType + restaurant + tone_prob, family=gaussian(), data=data.ah)
    print(paste(animal_arg, hemi_arg))
    print(summary(glm.ah))
    
    df <- reset_index(as.data.frame(coef(summary(glm.ah))[, c(1, 2, 4)]))
    df$animal = animal_arg
    df$hemi = hemi_arg
    df_list[[i]] = df
    i = i + 1
  }
}

dt_p_df <- do.call(rbind, df_list)
names(dt_p_df)[3:4] = c("stdE", "p_val")
dt_p_df %>% filter(p_val <= 0.05, grepl("stim", index))
```


```{r}
ivw_mean <- function(x, group) {
  v0 = var(x)
  v <- numeric(length(x))
  for (g in unique(group)) {
    v[group==g] = 1 / var(x[group == g])
  }
  return(weighted.mean(x, v))
}
ivw_mean(data$rt, data$uniqueid)

```

## Collapsing restaurant viz stim effect

```{r data_viz_stim, echo = TRUE, message=FALSE, warning=FALSE, fig.width=10, fig.height=6}
df_list = list()


# data %>% select(animal, restaurant, tone_prob, stimType, accept, plugged_in, quit_time) %>%
#   filter(stimType %in% c('stim', 'nostim'), is.na(quit_time)) %>%
#   mutate(tone_prob=as.factor(tone_prob), stimType=as.factor(stimType)) %>% 
#   group_by(animal, plugged_in, tone_prob, stimType) %>% 
#   summarise(accept_rate = mean(accept), accept_sd = sqrt(var(accept)), n=n()) %>% 
#   mutate(accept_se=accept_sd / sqrt(n)) %>%
#   
#   ggplot(aes(x=tone_prob, y=accept_rate, color=stimType)) +
#   geom_bar(stat="identity", alpha=0.5) +
#   geom_errorbar( aes(x=Species, ymin=mean-se, ymax=mean+se), width=0.4, colour="orange",
#   geom_boxplot() +
#   facet_grid(animal~plugged_in) +
#   theme_classic()

data %>% select(animal, restaurant, tone_prob, stimType, accept, plugged_in, quit_time) %>%
  filter(stimType %in% c('stim', 'nostim'), is.na(quit_time)) %>%
  mutate(tone_prob=as.factor(tone_prob), stimType=as.factor(stimType)) %>% 
  group_by(animal, plugged_in, tone_prob, stimType) %>% 
  ggplot(aes(x=tone_prob, y=accept, color=stimType)) +
  stat_summary(fun = mean,
               geom = "pointrange",
               fun.max = function(x) mean(x) + sd(x) / sqrt(length(x)),
               fun.min = function(x) mean(x) - sd(x) / sqrt(length(x))) +
  facet_grid(plugged_in~animal)

data %>% select(animal, restaurant, tone_prob, stimType, accept, plugged_in, quit_time) %>%
  filter(plugged_in %in% c("L", "R"), is.na(quit_time)) %>%
  mutate(tone_prob=as.factor(tone_prob), stimType=as.factor(stimType)) %>% 
  group_by(animal, plugged_in, tone_prob, stimType) %>% 
  ggplot(aes(x=tone_prob, y=accept, color=stimType)) +
  stat_summary(fun = mean,
               geom = "pointrange",
               fun.max = function(x) mean(x) + sd(x) / sqrt(length(x)),
               fun.min = function(x) mean(x) - sd(x) / sqrt(length(x))) +
  facet_grid(plugged_in~animal)

```

```{r stim_only_summary_allR, echo = TRUE, message=FALSE, warning=FALSE}
df_list = list()
i = 1
for (animal_num in c(39:40, 42:45)) {
  animal_arg <- sprintf("RRM0%d", animal_num)
  for (hemi_arg in c("L", "R")) {
    data.ah <- data %>% 
    select(animal, restaurant, tone_prob, stimType, accept, plugged_in, quit_time) %>% 
    filter(stimType %in% c('stim', 'nostim'), animal==animal_arg, plugged_in==hemi_arg, is.na(quit_time))
    glm.ah <- glm(accept ~ tone_prob:stimType + stimType + tone_prob, family=binomial(), data=data.ah)
    print(paste(animal_arg, hemi_arg))
    print(summary(glm.ah))
    
    df <- reset_index(as.data.frame(coef(summary(glm.ah))[, c(1, 2, 4)]))
    df$animal = animal_arg
    df$hemi = hemi_arg
    df_list[[i]] = df
    i = i + 1
  }
}

accept_p_df <- do.call(rbind, df_list)
names(accept_p_df)[3:4] = c("stdE", "p_val")
accept_p_df %>% filter(p_val <= 0.05, grepl("stim", index))
```


```
toneprob +  restaurant*animal
```

